Linked Open Social Data for Scientific Benchmarking

### 2016-07-21 paper.tex
advances in query formating

\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{changepage}
\usepackage{longtable}
\usepackage{tabularx}
% \usepackage[showframe=true]{geometry}
%
\newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
\pgfplotsset{compat=1.5}
\pgfplotsset
{
	width=0.5\textwidth,
	x tick label style={/pgf/number format/1000 sep=},
  enlarge x limits = 0.0,
  ymajorgrids=true,
	major tick style={draw=none},
  ymin = 0.0,
	every axis/.append style={
		every x tick label/.append style={font=\tiny},
    every y tick label/.append style={font=\tiny},
    every axis label/.append style={font=\small},
    height=37mm,
    width=37mm,
    title style={at={(0.5,0.90)}, font=\normalfont},
    xticklabel style={yshift=4pt}
	}
}

@@

We developed an ontology synthesis method that probes the ontological structure in data with
SPARQL queries and post-processing which can be divided in the following steps:
\begin{enumerate}[leftmargin=0cm]
    \item Obtaining all distinct classes with the query:\\
        \textttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
    \item Obtaining all distinct properties with the query:\\
        \textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
    \item For each class, get distinct subject classes and predicates where the
        the object is an instance of the class:\\
        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
        a ?cs . \}}
    \item For each class, get distinct predicates and object classes or
        datatypes where the subject is an instance of such class:\\
        \textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
                a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \}}
    \item For each property, check if it is functional, i.e. if it
        occurs only once with each subject:\\
        \textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
            <property\_uri> ?o \} GROUP BY ?s}
    \item For each property, find the incident range and domain with the
        queries:\\
        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}} \\
        and \\
        \textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
    \item For each instance of each class, get all distinct predicates.
        For each predicate, check if all instances of the class
        hold such relationship (existential restriction):\\
        \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
        \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
        \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
                a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
    \item and if all instances that hold such relationship are instances of the class
        (universal restriction):\\
        \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
    \item Draw each class, each property and the overall figure.
    \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
        statements to better organize knowledge and link to third party
        ontologies and data.
\end{enumerate}

@@

\subsection{SPARQL queries}\label{queries}
There are numerous useful and general purpose SPARQL queries to be performed agaist the LOSD database.
Here we write some of the most basic of such queries selected by their potential to be varied.
All queries assume the use the preable \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
\begin{enumerate}[leftmargin=0cm]
	\item Retrieve the number of participants
	\item Retrieve the number of relations, be them interactions of friendships
	\item Retrieve the number of relations, be them interactions of friendships
	\item Retrieve all text produced by an specific user
	\item List 100 users with the most friendships
	\item List 100 users with most interactions
	\item Search string in messages
\end{enumerate}

>>>>

\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{float}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% \usepackage{changepage}
\usepackage{longtable}
\usepackage{tabularx}
% \usepackage[showframe=true]{geometry}
%
\newcommand{\textttt}[1] {\texttt{\footnotesize#1}}
\newcommand{\h} {\hphantom ~ }
% \newcommand{\textttt}[1] {\mbox{\texttt{\footnotesize#1}}}
% \newcommand{\textttt}[1] {
% \begin{verbatim} #1 \end{verbatim}
% }
\pgfplotsset{compat=1.5}
\pgfplotsset
{
	width=0.5\textwidth,
	x tick label style={/pgf/number format/1000 sep=},
  enlarge x limits = 0.0,
  ymajorgrids=true,
	major tick style={draw=none},
  ymin = 0.0,
	every axis/.append style={
		every x tick label/.append style={font=\tiny},
    every y tick label/.append style={font=\tiny},
    every axis label/.append style={font=\small},
    height=37mm,
    width=37mm,
    title style={at={(0.5,0.90)}, font=\normalfont},
    xticklabel style={yshift=4pt}
	}
}

@@

We developed an ontology synthesis method that probes the ontological structure in data with
SPARQL queries and post-processing.
The results are OWL code and figures which are available in the
Supporting Information file.
The method can be improved to comprise further OWL axioms,
but is currently performed to fit present needs as follows:
\begin{enumerate}[leftmargin=0cm]
    \item Obtaining all distinct classes with the query:\\
        % \textttt{SELECT DISTINCT ?class WHERE { ?s a ?class }}
        \textttt{SELECT DISTINCT ?class WHERE \{ ?s a ?class \}}
    \item Obtaining all distinct properties with the query:\\
        \textttt{SELECT DISTINCT ?p WHERE \{ ?s ?p ?o \}}
    \item For each class, get distinct subject classes and predicates where the
        the object is an instance of the class:\\
        \textttt{SELECT DISTINCT ?p ?cs WHERE \{ ?i a <class\_uri> . ?s ?p ?i . ?s
        a ?cs . \}}
    \item For each class, get distinct predicates and object classes or
        datatypes where the subject is an instance of such class:\\
        \textttt{SELECT DISTINCT ?p ?co (datatype(?o) as ?do) WHERE \{ ?i
                a <class\_uri> . ?i ?p ?o . OPTIONAL \{ ?o a ?co . \} \}}
    \item For each property, check if it is functional, i.e. if it
        occurs only once with each subject:\\
        \textttt{SELECT DISTINCT (COUNT(?o) as ?co) WHERE \{ ?s
            <property\_uri> ?o \} GROUP BY ?s}
    \item For each property, find the incident range and domain with the
        queries:\\
        \textttt{SELECT DISTINCT ?co (datatype(?o) as ?do) WHERE \{ ?s
                <property\_uri> ?o . OPTIONAL \{ ?o a ?co . \} \}} \\
        and \\
        \textttt{SELECT DISTINCT ?cs WHERE \{ ?s <property\_uri> ?o . ?s a ?cs . \}}
    \item For each instance of each class, get all distinct predicates.
        For each predicate, check if all instances of the class
        hold such relationship (existential restriction):\\
        \textttt{SELECT DISTINCT ?p WHERE \{ ?s a <class\_uri>. ?s ?p ?o . \}}\\
        \textttt{SELECT DISTINCT ?s WHERE \{ ?s a <class\_uri> \}}\\
        \textttt{SELECT DISTINCT ?s ?co  (datatype(?o) as ?do) WHERE \{?s
                a <class\_uri>. ?s <property\_uri> ?o . OPTIONAL \{?o a ?co . \}\}}
    \item and if all instances that hold such relationship are instances of the class
        (universal restriction):\\
        \textttt{SELECT DISTINCT ?s WHERE \{ ?s <property\_uri> ?o . \}}
    \item Draw each class, each property and the overall figure.
    \item Make \textttt{rdfs:subClassOf} and \texttt{rdfs:subPropertyOf}
        statements to better organize knowledge and link to third party
        ontologies and data.
\end{enumerate}

@@

\subsection{SPARQL queries}\label{queries}
There are numerous useful and general purpose SPARQL queries to be performed agaist the LOSD database.
Here we write some of the most basic of such queries selected by their potential to be varied.
All queries assume the use the preable \textttt{PREFIX po: <http://purl.org/socialparticipation/po/>}.
\begin{enumerate}[leftmargin=0cm]
	\item Retrieve the number of participants:\\
            \textttt{SELECT (COUNT(DISTINCT ?author) as ?c) WHERE \{
            ?author a po:Participant . \} }
	\item Retrieve the number of relations, be them interactions or
            friendships:\\
            \textttt{SELECT (COUNT(?interaction) as ?c) WHERE \{\\
                    \h \{ ?interaction a po:Friendship \} UNION \{ ?interaction
                    a po:Interaction \} UNION\\
                    \h \{ ?interaction po:retweetOf
                    ?message \} UNION \{ ?interaction po:replyTo ?message
                    \}\\
                    \h UNION \{ ?interaction po:directedTo ?participant
                    \}\\ \} }
	\item Retrieve all text produced by an specific user:\\
            \textttt{SELECT (CONCAT(?text) as ?texts) WHERE \{\\
                    \h ?activity po:author <user\_uri> . ?activity po:text ?text .\\
    \}}
        \item List 1000 users (URIs and names) with the most friendships and the number of
            friendships in descending order by the number of friendships:\\
            \textttt{SELECT DISTINCT ?participant (COUNT(?interaction)
            as ?c) WHERE \{\\
                \h ?interaction a po:Friendship . ?interaction po:member ?participant . \\
            \} ORDER BY DESC(?c) LIMIT 1000}
	\item Search string ``pineapple'' in LOSD messages:\\
            \textttt{SELECT ?text WHERE \{ \\
                    \h ?activity po:text ?text . FILTER regex(?text, 'pineapple', 'i')\\
    \}}
\end{enumerate}
